---
title: "Portfolio_assignment_7"
author: "Bas Koppenaal"
date: "13-5-2021"
output: html_document
---
1. Load the flu (“./data/flu_data.csv), the dengue (.”/data/dengue_data.csv) and the gapminder ({dslabs} package) into three separate dataframes in R.

Done

2. Check if they are in the right shape. Is the data in the ‘tidy’ format? If not change the format to ‘tidy’. 

I made flue_data and dengue_data tidy, Gapminder was already tidy.

3.Change the country and date variables of the three tables so that they coincide in terms of data type, class and values 
      A. all the Class of Country is Factor, Type is integer
      B. all the Class of date is Date, gapminder doesnt have date but has year. For now i leave gapminder$year as integer totdat ik hiervoor moet terugkeren
      C. Wat bedoelen ze met Values---------------------
      --------------------------------
      --------------------------------
      -----------------------------
4. Store the three tables as separate (so six in total) .csv and .rds files.
Done

5. In Dbeaver create a new PostgreSQL database “workflowsdb”
Done

6. Took me a while since i didnt understand the question right. I tried to get the CSV and RSD files into Docker instead of using R to get the files in. However the files have been put in Docker

7. I didnt know what to expect so i just did some exercises on the code.

8. Same as point 7 i didnt know what kind of data i wanted to collect 

9. 



```{r setup, include=FALSE}
library(dslabs)
library(readr)
library(tidyverse)
library(RPostgreSQL)
library(devtools)
library(remotes)
library(RPostgres)

```



```{r cars}



##load in flu_data.csv, skip the metadata
flu_data <- read_csv(here::here("Assignment7/flu_data.csv"), skip = 10)
##load in dengue_data.csv, skip the metadata 
dengue_data <- read_csv(here::here("Assignment7/dengue_data.csv"), skip = 10)

##flue_data not tidy
## View(flu_data)

##dengue_data not tidy
## View(dengue_data)

##gapminder data needs to be displayed different
## View(gapminder)
 
##making flu data tidy
Tidy_flu_data <- flu_data %>% 
  pivot_longer(c( "Argentina","Australia", "Austria" ,"Belgium" ,"Bolivia" ,"Brazil" ,"Bulgaria" ,"Canada", "Chile" ,"France" ,"Germany" ,"Hungary" ,"Japan" ,"Mexico" ,"Netherlands" ,"New Zealand" ,"Norway" ,"Paraguay" ,"Peru" ,"Poland" ,"Romania" ,"Russia" ,"South Africa", "Spain" ,"Sweden" ,"Switzerland" ,"Ukraine" ,"United States" , "Uruguay"),
               names_to = "country", values_to = "values") # combine col names using an underscore




##making data tidy
Tidy_dengue_data <- dengue_data %>% 
  pivot_longer(c( "Argentina" ,"Bolivia", "Brazil", "India", "Indonesia" , "Mexico" , "Philippines" ,"Singapore" ,"Thailand","Venezuela"),
               names_to = "country", values_to = "values") 


##making the class of Country in gapminder in a Character
gapminder$country <- as.character(gapminder$country)


Tidy_dengue_data$values <- as.numeric(Tidy_dengue_data$values)
Tidy_flu_data$values <- as.numeric(Tidy_flu_data$values)

##   opsplitsen datum in de vakken jaar dag en maand, soort datum toevoegen en maken genaamd flu_data en dengue_data



##   infant_mortality life_expectancy fertility population gdp  maken in een lijst van soort data en values


##Tidy flue data
View(Tidy_flu_data)

##Tidy dengue data
View(Tidy_dengue_data)

##Tidy gapminder
View(gapminder)

##save everything as csv
write.csv(Tidy_flu_data,"Assignment7\\Tidy_flu_data.csv", row.names = FALSE)
write.csv(Tidy_dengue_data,"Assignment7\\Tidy_dengue_data.csv", row.names = FALSE)
write.csv(gapminder,"Assignment7\\gapminder.csv", row.names = FALSE)

##and save everything as RDS
saveRDS(Tidy_flu_data, "Assignment7\\Tidy_flu_data.rds")
saveRDS(Tidy_dengue_data, "Assignment7\\Tidy_dengue_data.rds")
saveRDS(gapminder, "Assignment7\\gapminder.rds")

```


```{R Exporting data}
## 7.6
## add password if you want to run the file, but for security i wont put it in here. But this peace of code makes a connection to DOCKER.
con <- dbConnect(RPostgres::Postgres(), 
                 dbname = "workflowsdb", 
                 host="localhost", 
                 port="5432", 
                 user="postgres", 
                 password="") 

##looking at talbes in file
dbListTables(con) 

## putting the code in Docker
dbWriteTable(con, "Tidy_flu_data", Tidy_flu_data)
dbWriteTable(con, "Tidy_dengue_data", Tidy_dengue_data)
dbWriteTable(con, "gapminder", gapminder)

##7.7 I thould it might be easyr to put the code in here.

## SELECT
##     "Date",
## 	country

## FROM
##     public."Tidy_flu_data"
    
## ORDER BY
##     country desc,
##    	"Date" ASC;
    
   
##7.8  collecting some data from the file

Tidy_flu_data %>% dplyr::filter(Date=='2002-12-29') %>% select(country)


```

