---
title: "Portfolio_assignment_7"
author: "Bas Koppenaal"
date: "13-5-2021"
output: html_document
---
1. Load the flu (“./data/flu_data.csv), the dengue (.”/data/dengue_data.csv) and the gapminder ({dslabs} package) into three separate dataframes in R.

Done

2. Check if they are in the right shape. Is the data in the ‘tidy’ format? If not change the format to ‘tidy’. 

I made flue_data and dengue_data tidy, Gapminder was already tidy.

3.Change the country and date variables of the three tables so that they coincide in terms of data type, class and values 
      A. all the Class of Country is Factor, Type is integer
      B. all the Class of date is Date, gapminder doesnt have date but has year. For now i leave gapminder$year as integer totdat ik hiervoor moet terugkeren
      C. Wat bedoelen ze met Values---------------------
      --------------------------------
      --------------------------------
      -----------------------------
4. Store the three tables as separate (so six in total) .csv and .rds files.
Done

5. In Dbeaver create a new PostgreSQL database “workflowsdb”
Done

6. Took me a while since i didnt understand the question right. I tried to get the CSV and RSD files into Docker instead of using R to get the files in. However the files have been put in Docker

7. I didnt know what to expect so i just did some exercises on the code.

8. Same as point 7 i didnt know what kind of data i wanted to collect 

9. Took some time but i changed all the data frames: flue data and dengue data have their dates seperated so they have a collum called year. gapminder has all datas now in 2 collumns called data-name and values. also added everywhere ID and added named_data to flue and dengue data so if they are joined it is clear what is dengue data and what is flue data. 

10. i updated what i did in previous areas so the clean data is now loaded in.

11. i joined denque and flu data in R with dplr, if i need to join everything together or do it in dbeaver let me know but when i read the assignment this seems to be fine.

12. i first exported it to dbeaver and than back.


```{r setup, include=FALSE}
library(dslabs)
library(readr)
library(tidyverse)
library(RPostgreSQL)
library(devtools)
library(remotes)
library(RPostgres)

```



```{r cars}



##load in flu_data.csv, skip the metadata
flu_data <- read_csv(here::here("Assignment7/flu_data.csv"), skip = 10)
##load in dengue_data.csv, skip the metadata 
dengue_data <- read_csv(here::here("Assignment7/dengue_data.csv"), skip = 10)
##7.12 Loading back the data in
dengue_flue_data_dbeaver <- read_csv(here::here("Assignment7/Dengue_flue_data__202105241624.csv"))


##flue_data not tidy
## View(flu_data)

##dengue_data not tidy
## View(dengue_data)

##gapminder data needs to be displayed different
## View(gapminder)
 
##making flu data tidy
Tidy_flu_data <- flu_data %>% 
  pivot_longer(c( "Argentina","Australia", "Austria" ,"Belgium" ,"Bolivia" ,"Brazil" ,"Bulgaria" ,"Canada", "Chile" ,"France" ,"Germany" ,"Hungary" ,"Japan" ,"Mexico" ,"Netherlands" ,"New Zealand" ,"Norway" ,"Paraguay" ,"Peru" ,"Poland" ,"Romania" ,"Russia" ,"South Africa", "Spain" ,"Sweden" ,"Switzerland" ,"Ukraine" ,"United States" , "Uruguay"),
               names_to = "country", values_to = "Flue_data") # combine col names using an underscore

##making data tidy
Tidy_dengue_data <- dengue_data %>% 
  pivot_longer(c( "Argentina" ,"Bolivia", "Brazil", "India", "Indonesia" , "Mexico" , "Philippines" ,"Singapore" ,"Thailand","Venezuela"),
               names_to = "country", values_to = "Dengue_data") 


##making the class of Country in gapminder in a Character
gapminder$country <- as.character(gapminder$country)

## values nummeriek maken -- Of wil ik dit als float hebben?
Tidy_dengue_data$Dengue_data <- as.numeric(Tidy_dengue_data$Dengue_data)
Tidy_flu_data$Flue_data <- as.numeric(Tidy_flu_data$Flue_data)

##   opsplitsen datum in de vakken jaar dag en maand, soort datum toevoegen en maken genaamd flu_data en dengue_data. Dit wordt gedaan aangezien gapminder alleen jaar heeft en nu kunnen ze mergen. Mischien beter van niet

Tidy_flu_data <- separate(Tidy_flu_data, "Date", c("year", "month", "day"), sep = "-")
Tidy_dengue_data <- separate(Tidy_dengue_data, "Date", c("year", "month", "day"), sep = "-")


## give each measured data a number. Does Gapminder also need a number?  might need this later. Or do we even need ID at all?

##Tidy_flue_data_ID <- Tidy_flu_data %>%  mutate(ID=seq(nrow(.)))
##Tidy_dengue_data_ID <- Tidy_dengue_data %>%  mutate(ID=seq(nrow(.)))
##gapminder_ID <- gapminder %>%  mutate(ID=seq(nrow(.)))

##   infant_mortality life_expectancy fertility population gdp  maken in een lijst van soort data en values

gapminder_tidy <- gapminder_ID %>%
  pivot_longer(cols = c(`infant_mortality`, `life_expectancy`, 'fertility', 'population', 'gdp'), names_to = "Data_name", values_to = "values")

## adding Data_name to dengue data and flue data so they can be joined together
##Tidy_dengue_data_ID_Named <- Tidy_flue_data_ID %>%  mutate(Data_name= 'dengue_data')
##Tidy_flue_data_ID_Named <- Tidy_dengue_data_ID %>%  mutate(Data_name= 'flue_data')

## 7.11 Here i join dengue and flue data together, Could be done with leftjoin or merge
## Dengue_flue_data <- merge(Tidy_flu_data , Tidy_dengue_data ,by= c("year", "month", "day", "country") ) 
 Dengue_flue_data <-left_join(Tidy_flu_data, Tidy_dengue_data, by = c("year", "month", "day", "country"))


## to make it tidy
 Tidy_Dengue_flue_data <- Dengue_flue_data %>% 
  pivot_longer(c(Flue_data , Dengue_data ),
               names_to = "Data_name", values_to = "values") 
 
gapminder_tidy <- gapminder_ID %>%
  pivot_longer(cols = c(`infant_mortality`, `life_expectancy`, 'fertility', 'population', 'gdp'), names_to = "Data_name", values_to = "values")

##Tidy flue data
View(Tidy_dengue_data_ID_Named)

##Tidy dengue data
View(Tidy_flue_data_ID_Named)

##Tidy gapminder
View(gapminder_tidy)

##save everything as csv  Needs to be updated at the end
write.csv(Tidy_flu_data,"Assignment7\\Tidy_flu_data.csv", row.names = FALSE)
write.csv(Tidy_dengue_data,"Assignment7\\Tidy_dengue_data.csv", row.names = FALSE)
write.csv(gapminder,"Assignment7\\gapminder.csv", row.names = FALSE)

##and save everything as RDS needs to be updated at the end
saveRDS(Tidy_flu_data, "Assignment7\\Tidy_flu_data.rds")
saveRDS(Tidy_dengue_data, "Assignment7\\Tidy_dengue_data.rds")
saveRDS(gapminder, "Assignment7\\gapminder.rds")

```


```{R Exporting data}
## 7.6
## add password if you want to run the file, but for security i wont put it in here. But this peace of code makes a connection to DOCKER.
con <- dbConnect(RPostgres::Postgres(), 
                 dbname = "workflowsdb", 
                 host="localhost", 
                 port="5432", 
                 user="postgres", 
                 password="") 

##looking at talbes in file
dbListTables(con) 

## putting the code in Docker
dbWriteTable(con, "flue_data", Tidy_flue_data_ID_Named)
dbWriteTable(con, "dengue_data", Tidy_dengue_data_ID_Named)
dbWriteTable(con, "gapminder", gapminder_tidy)
dbWriteTable(con, "Dengue_flue_data", Tidy_Dengue_flue_data)

##7.7 I thould it might be easyr to put the code in here.

## SELECT
##     "year",
## 	country

## FROM
##     public."flu_data"
    
## ORDER BY
##     country desc,
##    	"year" ASC;
    
   
##7.8  collecting some data from the file

Tidy_flu_data %>% dplyr::filter(Date=='2002-12-29') %>% select(country)


```

